\section{Introduction}

In \emph{The Black Sheep}, Calvino begins from a form of order that is plainly criminal and strangely balanced at the same time. Everyone steals at night and everyone is robbed in turn, so each household loses what another household gains and the society remains equal while living under universal predation. The story becomes economically acute when one man decides to remain at home. His abstention removes a link from the nightly circuit and disrupts the accounting on which the circuit rested: the burglar who intended to rob him returns empty handed, while another household, which would otherwise have been robbed by that failed burglar, keeps what it would have lost. The immediate consequence of honesty is therefore a local redistribution. Calvino then follows that redistribution toward a harder order in which some households become rich, poor households are recruited into theft on behalf of the rich, and later the rich hire guards and police to defend fortunes accumulated in the earlier phase \citep{calvino_1993}. The sequence is literary, but it poses an unusually clean economic problem.

The sequence admits a formal treatment of a regime of reciprocal predation that can reproduce itself as long as participation remains universal, but its initial balance requires exact conditions because a society of thieves remains equal only if every household is both predator and prey under the same accounting. Once one agent abstains, the relevant issue becomes the continuation incentives disturbed by that local break and the path by which reciprocity gives way to a different organization of predation. The later transition raises a further problem, because decentralized theft, delegated theft, and organized protection do not rest on the same primitives. The benchmark that isolates reciprocal balance cannot by itself generate class formation, while the environment that generates class formation obscures the balance from which the sequence begins.

Reciprocal balance is easiest to see in a unit transfer benchmark in which every successful raid moves one unit, every household begins with the same wealth, and a balanced cycle is a permutation in which every agent has one outgoing raid and one incoming raid. In that environment wealth is preserved period by period, and any such cycle is a weak Nash equilibrium under symmetry. The benchmark remains intentionally narrow because it does not yet contain the ingredients needed for hierarchy or enforcement; its role is to isolate the exact form of reciprocity that Calvino places at the start of the story and to give precise content to the claim that universal participation can sustain equality even inside a predatory order.

The first substantive break comes with the honest perturbation, for which the relevant equilibrium object is a stationary Markov regime in the repeated benchmark rather than a one period game or the full set of repeated outcomes supported by elaborate punishments. One agent is permanently honest and remains at home every period. All other agents choose whether to steal personally or remain at home as a function of current wealth. The first honest night creates a local windfall because one intended victim is no longer robbed, while the predecessor who targeted the honest house fails to obtain loot. Once leaving home exposes wealth left at home, that windfall changes at least one adjacent continuation payoff. Under a transparent wealth threshold condition, the original pattern of balanced reciprocity cannot remain stationary with exactly one permanently honest agent.

Calvino carries the sequence well beyond the first rupture, because a society organized around universal corruption does not move from corruption to honesty in one step but first toward stratified predation and only later toward selective protection. Making that transition economically coherent requires an explanation of why richer households stop stealing personally, why poorer households accept predatory work on behalf of the rich, and why those same elites later prefer guards to thieves. The proportional transfer extension supplies those margins. When theft scales with the wealth of the victim, a rich household that goes out to steal exposes a larger stock at home. Delegation can then dominate personal theft if delegated predation has a targeting advantage and if poor workers face weak outside options. Patron surplus is simply the residual between delegated gross returns and the outside option of the worker net of contracting cost. Later, once elite wealth becomes large and the prey base among the poor is depleted, the value of protection overtakes the value of one more round of delegated predation. The chronology then moves from reciprocal theft to patronage and from patronage to guarding for reasons that are explicit in the payoff comparisons.

Corruption equilibria remain privately coherent when individual choices change the environment faced by others \citep{andvig_moene_1990,bardhan_1997,shleifer_vishny_1993,tirole_1996}; conventions and equilibrium selection explain how local perturbations move societies across self reproducing orders \citep{young_1993}; appropriation and protection allocate effort between predation and defense under insecure claims \citep{skaperdas_1992,grossman_kim_1995,hirshleifer_1995}; hierarchy and delegated organization explain how vertical control redistributes rents \citep{olsen_torsvik_1998}; and endogenous institutions explain why organized enforcement often protects concentrated advantage \citep{olson_1993,north_wallis_weingast_2009,besley_persson_2009,acemoglu_ticchi_vindigni_2011}. The novelty here lies in placing those margins inside one transition from balanced reciprocal theft to patronage and then to selective protection.

The analogy to top trading cycles needs careful handling because it is useful and easy to overstate. A balanced theft cycle resembles a cyclic allocation graph in one limited respect: each agent points outward and is pointed to by exactly one other agent. That graph structure helps explain why the benchmark preserves equality under symmetry and why a local break has consequences beyond the abstainer himself. But the analogy stops at the graph. In the Shapley and Scarf environment, transfers are voluntary and stability is defined through blocking coalitions \citep{shapley_scarf_1974}. In the present environment, theft is coercive, generates externalities, and changes later opportunities through exposure and depletion. The cyclic analogy therefore serves only as a disciplined description of structure. Claims about the core or about mechanism design do not survive the change in environment.

Because the benchmark conserves aggregate wealth by construction, aggregate wealth alone cannot distinguish regimes. Period welfare is therefore defined as the sum of concave utilities over wealth minus theft effort costs, contracting costs, guarding costs, and expected punishment losses. This object separates margins that would otherwise be mixed together: pure transfers need not change total wealth, the same transfers can lower utilitarian welfare once they produce inequality, and the later movement from reciprocal theft to patronage and guarding can lower welfare further because predation and protection consume resources.

The analytical results leave open questions of timing, prevalence, regime duration, and comparative statics that are difficult to settle in closed form once many heterogeneous agents interact, so the same payoff comparisons derived in the theory are embedded in an explicit transition process and studied in simulation. The resulting paths reproduce the first honest perturbation in event time, trace class formation in the proportional environment, map parameter regions associated with patronage, coexistence, guarding, and collapse, and quantify the evolution of inequality and welfare along the transition.

Two restrictions structure the analysis from the outset. No full Markov perfect equilibrium is solved for the large stochastic state space generated by delegated theft, matching frictions, and protection choices,\footnote{This division of labor is deliberate. The theory identifies the margins and thresholds that matter. The simulation studies how those margins interact once the full state space becomes too large for a complete analytical characterization.} and production is left outside the model. Wealth is redistributed, dissipated through cost, and protected without being endogenously created, which usefully isolates the political economy of predation but also means that the final guard regime should be read as protection of accumulated spoils rather than as a complete productive property-rights order.

The paper proceeds as follows. Section \ref{sec:literature} places the argument in the literatures on corruption, conventions, hierarchy, appropriation, and endogenous enforcement. Section \ref{sec:model} presents the unified model, beginning with the unit transfer benchmark and then moving to the repeated perturbation result, the patronage contract problem, the guarding threshold, and the welfare criterion. Section \ref{sec:simulation} translates those analytical objects into a reproducible simulation design with explicit timing, matching, and regime classification. Section \ref{sec:results} presents the figures and tables in the order required by the argument, from the first honest night to the later regime map and welfare decomposition. Section \ref{sec:discussion} closes by drawing the discussion and conclusion together before the bibliography and appendix.
