\section{Related Literature}
\label{sec:literature}

The problem posed by Calvino belongs to a familiar concern in economics, namely the possibility that conduct which is plainly undesirable in welfare terms can nevertheless form a coherent and durable equilibrium. In the corruption literature this concern appears in several adjacent forms. \citet{andvig_moene_1990} show how corruption can reproduce itself by altering the incentives faced by later actors; once others are expected to steal or accept bribes, honest conduct loses much of its private return. \citet{bardhan_1997} surveys the same logic at a broader level and stresses the institutional complementarities that make corrupt systems persistent. \citet{shleifer_vishny_1993} add the crucial organizational point that the social effects of corruption depend on whether predation is fragmented or coordinated, while \citet{tirole_1996} show how collective reputations can trap organizations in persistent patterns of misconduct even when many participants would prefer a cleaner order. The same concern appears here in a narrower and more sequential setting. Instead of bribery or bureaucratic malfeasance, the object is a world of reciprocal thieves and the transition set off by a small honest rupture that disturbs the accounting on which reciprocity depended.

That emphasis on disturbance and succession also places the analysis near the literature on conventions and equilibrium selection. \citet{young_1993} studies how recurrent interaction and local perturbations can move a society from one sustained convention to another. The relevant insight is that a local change can shift the continuation incentives of neighboring actors and thereby alter the basin of attraction of the entire system. The opening equilibrium in Calvino is a convention in exactly that sense. Its status is behavioral. Normative approval has nothing to do with its persistence. Under symmetry it is regular and reproducible, even though it rests on universal predation. The first honest agent interrupts a pattern of mutual dependence. The first windfall and the first shortfall then alter who still finds participation privately worthwhile. Corruption and equilibrium selection meet at that transition.

The movement from reciprocal predation to hierarchical predation requires a different body of work. A large part of the corruption literature begins with an organization that is already stratified and then studies collusion, monitoring, or internal incentive design. \citet{olsen_torsvik_1998} remain particularly useful here because they show that hierarchy can alter the distribution of rents and the feasibility of corruption itself. The same problem arises once rich households stop stealing personally and begin to buy predatory labor from poor households. The difference is that hierarchy is not assumed at the start. It has to be generated from the breakdown of a decentralized arrangement. Patronage therefore requires a strict surplus condition. Delegation cannot rest on vague appeals to elite power alone. It must be profitable because a patron directs theft more effectively than a freelancer can and because the reservation utility of poor workers remains compressed. The central question is how a vertical structure of predation is built out of what had previously been a horizontal order.

\citet{skaperdas_1992} and \citet{grossman_kim_1995} study environments in which insecure claims induce agents to allocate resources between conflict and defense. \citet{hirshleifer_1995} sharpens the same theme by analyzing the conditions under which anarchy can unravel and more selective forms of order emerge. \citet{demsetz_1967} remains a natural background reference because he frames the emergence of property rights as a response to rising gains from internalization. The connection is close in its treatment of appropriation and protection as alternative uses of resources, but the point of departure matters. The starting point is not a blank condition of disorder. It is an order in which insecurity of claims is universal, reciprocal, and roughly egalitarian. The later rise of guarding is therefore not the first appearance of order. One order replaces another, and the later order protects a distribution created during the transition itself.

The connection to state formation and endogenous enforcement is therefore especially important. \citet{olson_1993} distinguishes roving bandits from stationary ones and argues that an encompassing interest in future extraction can create incentives to provide order. \citet{north_wallis_weingast_2009} generalize that intuition into a theory of violence orders organized around elite management of rents. \citet{besley_persson_2009} study the conditions under which legal and fiscal capacity emerge, while \citet{acemoglu_ticchi_vindigni_2011} show that inefficient coercive states can persist once they serve the interests of those who control them. \citet{acemoglu_robinson_2008} likewise stress that institutional change often preserves elite power while leaving that power intact. The final regime belongs in that lineage. Guards and police do not arise from civic virtue or from a neutral concern with equal rights. They arise because concentrated wealth becomes worth defending and because labor can be reallocated from predation to protection without changing who controls the terms of social order. The fit with endogenous-institutions work lies there, even though the model remains much more stripped down than the historical theories just cited.

Once guarding enters the picture, \citet{becker_1968} and \citet{becker_stigler_1974} become relevant because elite households no longer face only the choice between stealing and abstaining. No benevolent state chooses optimal sanctions here, and no optimal policing problem is solved, but the comparison does shift toward two costly technologies, one extractive and one protective, each with its own labor demand and effect on future incentives.

Directed cycles matter here only as a graph theoretic analogy. A balanced theft cycle is also a directed cycle with one incoming and one outgoing link for each agent, and in \citet{shapley_scarf_1974} directed cycles organize voluntary exchange among agents who reassign indivisible objects. The analogy is useful at that level of structure and no further. The transfers here are coercive, alter later incentives through exposure and depletion, and impose externalities on actors who have not consented to any reassignment, so claims about the core, strategy proofness, or voluntary stability do not survive the change in environment.

\citet{skaperdas_1992} and \citet{grossman_kim_1995} study how insecure claims divert resources into appropriation and defense. \citet{becker_1968} and \citet{becker_stigler_1974} keep attention on sanction, precaution, and the incentives of enforcement. \citet{olson_1993}, \citet{north_wallis_weingast_2009}, \citet{besley_persson_2009}, and \citet{acemoglu_ticchi_vindigni_2011} show how organized coercion can stabilize extraction and protect concentrated advantage. Their intersection is the natural place for a setting that begins from egalitarian reciprocal corruption, because aggregate wealth cannot serve as the relevant welfare object once the benchmark preserves total wealth by construction. The comparison has to run through concavity in wealth together with the costs of theft, punishment, contracting, and guarding. Reciprocal corruption, hierarchical predation, and selective protection then appear as successive forms of one order, linked by a local honest rupture that shifts incentives without generating an impartial institutional exit.
