\section{Model}
\label{sec:model}

\subsection{The unit transfer raiding game}

The analysis begins from a repeated raiding game whose only purpose is to make the opening reciprocal order exact. Time is discrete and indexed by $t = 0,1,2,\dots$. Agents are indexed by $i \in \{1,\dots,N\}$ with $N \ge 3$, and wealth at the beginning of period $t$ is the publicly observed vector $w_t = (w_{1t},\dots,w_{Nt})$. In the benchmark all agents begin with common wealth $\bar{w} > \tau$, where $\tau > 0$ is the unit amount that can be stolen from an unoccupied house in one period. The richer proportional environment appears only after the benchmark and keeps the same state variable. The notation used throughout the paper is collected in Table \ref{tab:notation}; the same symbols are used in the theory, the simulations, and the appendix proofs.

\begin{table}[tbp]
    \centering
    \caption{Notation. The table collects the state variables, contract terms, and regime objects used throughout the theory and simulation sections.}
    \label{tab:notation}
    \input{tables/table_notation.tex}
\end{table}

In each benchmark period, agent $i$ chooses one action from the set
\[
A_i = \{H\} \cup \left(\{1,\dots,N\}\setminus\{i\}\right),
\]
where $H$ means staying home and $j \neq i$ means leaving home and attempting to raid household $j$. Occupancy determines vulnerability. If $a_{jt}=H$, household $j$ is occupied and no raid on $j$ succeeds in period $t$. If $a_{jt}\neq H$, household $j$ is unoccupied and can be raided. This is the only occupancy rule needed for the benchmark, and it makes the first honest night deterministic once one household stays home permanently.

Target congestion is explicit. For any household $j$, let
\[
M_{jt} = \{i \neq j : a_{it}=j\},
\qquad
m_{jt} = |M_{jt}|.
\]
If $a_{jt}=H$, every attempted raid on $j$ fails. If $a_{jt}\neq H$ and $m_{jt}\ge 1$, at most one unit $\tau$ can be extracted from $j$ during period $t$, and that unit is divided in expected value equally across the raiders in $M_{jt}$. The benchmark therefore treats loot as rival and rules out the possibility that several agents can each take the same unit independently.

Define the realized loss of household $i$ and the expected loot of agent $i$ by
\begin{equation}
\ell_{it}(a_t)
=
\tau \mathbf{1}\{a_{it}\neq H\}\mathbf{1}\{m_{it}\ge 1\},
\qquad
g_{it}(a_t)
=
\sum_{j\neq i}\mathbf{1}\{a_{it}=j\}\mathbf{1}\{a_{jt}\neq H\}\frac{\tau}{m_{jt}}.
\label{eq:loss-gain}
\end{equation}
The benchmark abstracts from direct theft cost so that the reciprocal accounting is exact. End-of-period wealth is therefore
\begin{equation}
w_{i,t+1} = w_{it} - \ell_{it}(a_t) + g_{it}(a_t).
\label{eq:benchmark-wealth-update}
\end{equation}

\subsection{Balanced reciprocity}

\begin{definition}[Balanced theft cycle]
\label{def:balanced-cycle}
A balanced theft cycle is a permutation $\sigma$ of $\{1,\dots,N\}$ with no fixed points such that each agent $i$ chooses $a_i=\sigma(i)$.
\end{definition}

Under a balanced theft cycle every household is unoccupied, every household is targeted by exactly one raider, and no target congestion appears because $m_j=1$ for every $j$. The basic accounting is then immediate but worth stating explicitly because the rest of the paper measures departures from this object.

\begin{lemma}[Balanced accounting]
\label{lem:balanced-accounting}
If all agents begin a period with common wealth $\bar{w}$ and play a balanced theft cycle, then every agent ends the period with wealth $\bar{w}$ and aggregate wealth is preserved.
\end{lemma}

\begin{proof}
See Appendix \ref{app:proofs}. Each agent loses exactly $\tau$ and gains exactly $\tau$, so wealth is unchanged agent by agent and therefore in aggregate.
\end{proof}

\begin{proposition}[Balanced reciprocity as a weak equilibrium]
\label{prop:balanced-cycle}
In the one-shot unit transfer stage game, any balanced theft cycle is a weak Nash equilibrium.
\end{proposition}

\begin{proof}
See Appendix \ref{app:proofs}. Deviating to staying home preserves the deviator's wealth but does not raise it, because the avoided loss is exactly offset by forgone loot. Deviating to a different target creates congestion, so expected loot falls below $\tau$ while the incoming loss remains. Hence no unilateral deviation is strictly profitable.
\end{proof}

Only a weak claim is needed at this stage. The purpose of the benchmark is not to establish efficiency or to characterize the full set of equilibria. It is to show that a reciprocal predatory order can be a coherent strategic object once actions and congestion are written down exactly. The resulting graph is a directed cycle, and that graph theoretic resemblance to top trading cycles remains useful when one link later breaks. The resemblance goes no further. Transfers here are coercive, occupancy changes the opportunity set of others, and rivalry in extraction creates externalities that have no analogue in the voluntary exchange environment of \citet{shapley_scarf_1974}.

\subsection{A permanent honest household and stationary continuation}

The repeated benchmark keeps the same stage game and adds two restrictions immediately. The horizon is infinite with common discount factor $\delta \in (0,1)$, and strategies are restricted to be stationary and Markov in current wealth. A strategy for agent $i$ is therefore a map from the current state $w_t$ into an action in $A_i$. This is the equilibrium class studied in the theorem below. The argument does not characterize the full set of repeated outcomes that could be sustained by history-dependent punishments.

Fix one agent $h$ who is permanently honest and therefore chooses $H$ in every period. Let $\sigma$ denote the pre-perturbation balanced theft cycle, let $p = \sigma^{-1}(h)$ be the predecessor who planned to rob $h$, and let $q = \sigma(h)$ be the successor whom $h$ would have robbed under universal participation. If every other agent continues the old cycle on the first perturbed date, then the occupancy rule makes the first wealth change deterministic:
\begin{equation}
w_{h1} = \bar{w}, \qquad w_{p1} = \bar{w} - \tau, \qquad w_{q1} = \bar{w} + \tau,
\label{eq:first-night-accounting}
\end{equation}
and all other agents remain at $\bar{w}$.

\begin{lemma}[First night local asymmetry]
\label{lem:first-night}
Under the benchmark cycle with one permanently honest agent $h$, the first honest night creates a local three agent asymmetry given by \eqref{eq:first-night-accounting}. In particular, the successor $q$ receives a windfall and the predecessor $p$ suffers an immediate shortfall.
\end{lemma}

\begin{proof}
See Appendix \ref{app:proofs}. The honest agent neither steals nor is robbed, the predecessor fails in his attempted raid on an occupied house, and the successor is not robbed by the honest agent.
\end{proof}

The benchmark stage game is enough to show that exact reciprocity breaks. A withdrawal result needs one additional margin. Suppose that personal raiding in the repeated game carries direct effort cost $c_T \ge 0$ and expected exposure loss $\rho w_{it}$, with $\rho \in (0,1)$, because the raider leaves current wealth unattended while away from home. The expected one-period return from personal theft is then
\begin{equation}
\pi^{S}(w_{it}) = \tau - \rho w_{it} - c_T,
\label{eq:self-theft-threshold}
\end{equation}
while the normalized return from staying home is $\pi^{N}(w_{it}) = 0$. Hence there is a wealth threshold
\begin{equation}
\widehat{w} = \frac{\tau - c_T}{\rho}
\label{eq:threshold}
\end{equation}
such that an agent weakly prefers personal theft when $w_{it} \le \widehat{w}$ and strictly prefers staying home when $w_{it} > \widehat{w}$.

\begin{assumption}[Interior exposure threshold]
\label{ass:threshold}
The exposure threshold satisfies $\bar{w} < \widehat{w} < \bar{w} + \tau$.
\end{assumption}

The assumption places the symmetric benchmark wealth level below the exposure cutoff and the first-night windfall above it. Personal theft remains privately viable at $\bar{w}$ but ceases to be optimal after one unit of windfall. This is a narrow comparative static statement and not a general theory of criminal choice.

\begin{theorem}[No stationary one-honest continuation]
\label{thm:no-stationary-one-honest}
Under Assumption \ref{ass:threshold}, there is no stationary Markov equilibrium of the repeated benchmark game that preserves the original balanced theft pattern for all agents other than the permanently honest agent $h$.
\end{theorem}

\begin{proof}
See Appendix \ref{app:proofs}. By Lemma \ref{lem:first-night}, the successor $q$ enters period $t=1$ with wealth $w_{q1} = \bar{w} + \tau$. Assumption \ref{ass:threshold} implies $w_{q1} > \widehat{w}$, so \eqref{eq:self-theft-threshold} yields $\pi^{S}(w_{q1}) < 0 = \pi^{N}(w_{q1})$. Thus $q$ strictly prefers staying home at $t=1$ and no longer continues to steal. Since the original balanced pattern requires $q$ to keep stealing, that pattern cannot be stationary after the perturbation.
\end{proof}

The theorem is deliberately local. It does not solve the full repeated game with one honest agent. It says only that the original balanced cycle cannot survive as a stationary Markov continuation once the first honest night moves the successor across the exposure threshold. That is exactly the comparison later implemented in the deterministic benchmark simulation.

Chronology matters. The honest agent is not robbed immediately upon becoming honest because occupied houses are perfectly safe in the unit transfer benchmark. Only later, after the paper moves to the proportional environment with delegated predation and guarding, do occupied but unguarded houses become imperfectly safe. The deterministic first-night result and the later stochastic vulnerability therefore belong to different layers of the model and should not be conflated.

\subsection{Proportional transfer extension and endogenous patronage}

After the first honest disturbance, wealth differences matter directly for choice because leaving home now exposes different stocks of wealth. The extension therefore keeps the same population, the same public wealth state, and the same logic of explicit target choice, while replacing the unit transfer rule with proportional appropriation. If household $j$ is successfully raided in period $t$, gross loot equals $\Gamma_{jt} = \gamma_j w_{jt}$, where $\gamma_j = \gamma \in (0,1)$ for nonelite households and $\gamma_j = \gamma_E \in (0,1)$ for elite households once the guarding margin becomes relevant. Rivalry in extraction remains explicit. For any household $j$, let $A_{jt}$ denote the set of active attackers who target $j$ in period $t$ and let $a_{jt} = |A_{jt}|$. Let $\omega_{jt}=1$ if household $j$ is occupied because its owner stays home, hires a thief, or hires a guard, and let $\omega_{jt}=0$ if the owner leaves home to raid personally or works away from home as a thief. Let $g_{jt}=1$ if household $j$ is guarded and zero otherwise. Conditional on being targeted, the probability that loot is successfully extracted from $j$ is
\begin{equation}
\kappa_{jt}
=
\mathbf{1}\{\omega_{jt}=0\}
+
\chi_t \mathbf{1}\{\omega_{jt}=1,\; g_{jt}=0\}
+
(1-q_t)\chi_t \mathbf{1}\{\omega_{jt}=1,\; g_{jt}=1\},
\label{eq:success-probability}
\end{equation}
where $\chi_t \in (0,1]$ is the vulnerability of an occupied but unguarded household and $q_t \in [0,1]$ is guard effectiveness. Expected gross loot taken from $j$ is therefore $\kappa_{jt}\Gamma_{jt}$, and if $a_{jt}\ge 1$ each attacker receives expected share $\kappa_{jt}\Gamma_{jt}/a_{jt}$. Congestion in the extension is thus handled in the same way as in the benchmark: several attackers can converge on one victim, but they divide one rival object.

Within each period, wealth $w_t$ is publicly observed before any sufficiently wealthy household chooses one occupation from
\[
A^{R}_t
=
\{\text{stay},\text{self raid},\text{offer theft contract},\text{offer guard job}\},
\]
while any poor household that receives offers chooses at most one action from
\[
A^{P}_t
=
\{\text{accept theft employment}, \text{accept guard employment}, \text{freelance raid}, \text{stay}\}.
\]
Every active thief next chooses a target, either directly if he is freelance or from the target class specified by the patron if he is employed, after which success, loot division, wages, remittances, and punishment are realized and wealth is updated. The patronage and guarding results below refer to this extension stage game and to stationary Markov behavior in its repeated version.

Two target classes keep the direction of predation clear. Let $T_t$ denote the set of vulnerable nonelite households available to a wealthy self raider, and let $H_t \subseteq T_t$ denote the richer and weakly protected part of that set that a patron can reach more effectively by directing a dependent thief. In the patronage regime the physical act of theft is therefore still aimed mainly at poor or near-poor households. The vertical transfer appears because contracted thieves remit part of the loot upward to the patron.

For a wealthy household that raids personally, let $m_t$ denote the expected gross return from the best target in $T_t$ under the success and congestion rule just stated. For the pairwise comparisons below it is enough to write
\begin{equation}
m_t = \gamma \bar{w}^{T}_t,
\label{eq:mt}
\end{equation}
where $\bar{w}^{T}_t$ is the mean wealth of the best available target class among nonelite households after accounting for the fact that concurrent raiders dilute one another's shares. If rich agent $i$ raids personally, the expected one period payoff is
\begin{equation}
\pi^{\mathrm{self}}_{it} = m_t - \rho w_{it} - c_T,
\label{eq:personal-payoff}
\end{equation}
where $\rho w_{it}$ is the expected value of leaving wealth exposed at home and $c_T$ is direct theft cost. Staying home without contracting yields the normalized payoff $\pi^{\mathrm{stay}}_{it}=0$.

Delegation alters both the target set and the division of returns. A theft contract offered by patron $i$ to poor worker $j$ specifies a fixed payment $s_{ijt}$, a thief loot share $\alpha_{ijt}\in[0,1]$, and access to the target class $H_t$. Let $\phi_t\ge 1$ summarize the patron's targeting advantage inside that class. Expected gross delegated loot is then
\begin{equation}
L_t = \phi_t \gamma \bar{w}^{H}_t,
\label{eq:delegated-loot}
\end{equation}
where $\bar{w}^{H}_t$ is mean wealth in $H_t$. If the worker rejects the offer, he can freelance on the remaining vulnerable poor. Let $o_{jt}$ denote that outside option:
\begin{equation}
o_{jt} = \lambda_t \gamma \bar{w}^{P}_t - c_T,
\qquad \lambda_t \in (0,1),
\label{eq:outside-option}
\end{equation}
where $\bar{w}^{P}_t$ is mean poor wealth and $\lambda_t$ summarizes the weaker targets and heavier congestion faced by freelancers. Worker $j$ accepts the patron's offer if
\begin{equation}
s_{ijt} + \alpha_{ijt}L_t \ge o_{jt}.
\label{eq:IR}
\end{equation}

Once a contract is accepted, one contracted theft relation is formed in that period. The patron pays contracting cost $k>0$, the worker raids within $H_t$, the patron remains occupied at home, and realized loot is divided according to the contract. Writing expected net joint surplus as
\begin{equation}
\Sigma_t = L_t - k - o_{jt},
\label{eq:total-surplus}
\end{equation}
and letting $\beta \in [0,1]$ denote patron bargaining power, the contract allocates
\begin{equation}
s_{ijt} + \alpha_{ijt}L_t = o_{jt} + (1-\beta)\Sigma_t
\label{eq:bargaining-rule}
\end{equation}
to the thief, leaving the patron with $\beta\Sigma_t$. The induced expected wealth update for the patron-worker pair is
\begin{equation}
\begin{aligned}
w_{i,t+1} &= w_{it} - \ell_{it} + (1-\alpha_{ijt})L_t - s_{ijt} - k,\\
w_{j,t+1} &= w_{jt} - \ell_{jt} + s_{ijt} + \alpha_{ijt}L_t - c_T,
\end{aligned}
\label{eq:contract-update}
\end{equation}
where $\ell_{it}$ and $\ell_{jt}$ denote realized incoming losses under the success and congestion rule in \eqref{eq:success-probability}. Rich agents who raid personally, poor freelancers, and guards are governed by the same accounting: period wealth equals initial wealth minus incoming loss plus role-specific income minus the resource cost attached to that role.

\begin{lemma}[Contract feasibility]
\label{lem:contract-feasibility}
There exists a patron-thief contract satisfying the thief's participation constraint if and only if
\begin{equation}
L_t - o_{jt} > k.
\label{eq:feasibility}
\end{equation}
\end{lemma}

\begin{proof}
See Appendix \ref{app:proofs}. A contract is feasible exactly when gross delegated loot exceeds the reservation utility of the thief plus the contracting cost of the patron.
\end{proof}

Condition \eqref{eq:feasibility} identifies the wedge that supports delegated predation. A patron gains nothing from hiring merely because wealth is high. Hiring becomes viable when access to $H_t$ raises expected gross loot above what a freelancer can secure elsewhere and when congestion keeps the worker's outside option below that delegated return. Without that wedge, the contract market has no surplus to divide.

\begin{proposition}[Strict patron surplus and occupational choice]
\label{prop:patronage}
Suppose Lemma \ref{lem:contract-feasibility} holds and the patron captures fraction $\beta$ of the contract surplus. Then patron $i$'s expected payoff from hiring thief $j$ is
\begin{equation}
\pi^{\mathrm{hire}}_{ijt} = \beta\left(L_t - k - o_{jt}\right).
\label{eq:patron-payoff}
\end{equation}
The rich agent strictly prefers hiring thief $j$ to both abstention and personal theft whenever
\begin{equation}
\beta\left(L_t - k - o_{jt}\right) > \max\left\{0,\; m_t - \rho w_{it} - c_T\right\}.
\label{eq:patron-dominance}
\end{equation}
\end{proposition}

\begin{proof}
See Appendix \ref{app:proofs}. Equation \eqref{eq:patron-payoff} follows directly from the bargaining rule. Hiring dominates abstention because the right-hand side of \eqref{eq:patron-dominance} exceeds zero, and it dominates personal theft because the expected surplus from delegation exceeds the net return available to the rich agent who exposes wealth at home.
\end{proof}

Surplus arises because delegation improves access to targets and because the worker's reservation payoff remains below delegated gross returns. Better information, denser local knowledge, asymmetric bargaining power, or labor-market slack can each sustain that wedge. The poor workers who accept these offers remain poor enough that leaving home exposes little wealth, while the compensation implied by \eqref{eq:bargaining-rule} still exceeds what freelance raiding would bring. The same pool of labor later becomes available for guarding.

\begin{corollary}[Comparative statics of patronage]
\label{cor:patronage-statics}
Fix $w_{it}$ and suppose $L_t = \phi_t \gamma \bar{w}^{H}_t$ and $o_{jt} = \lambda_t \gamma \bar{w}^{P}_t - c_T$. Whenever \eqref{eq:feasibility} holds, patron payoff \eqref{eq:patron-payoff} is increasing in $\phi_t$, decreasing in $\lambda_t$, decreasing in $k$, and decreasing in the attractiveness of personal theft only through the benchmark on the right-hand side of \eqref{eq:patron-dominance}. In particular, increases in delegation quality and decreases in poor outside options expand the set of states in which rich agents choose patronage over both self-theft and abstention.
\end{corollary}

\begin{proof}
See Appendix \ref{app:proofs}. The derivative of \eqref{eq:patron-payoff} with respect to $\phi_t$ is positive, the derivatives with respect to $\lambda_t$ and $k$ are negative, and the comparison against personal theft depends on whether these changes also widen the gap in \eqref{eq:patron-dominance}.
\end{proof}

Accordingly, patronage expands when delegated targeting becomes more effective, contracts when the outside option of poor workers strengthens, and weakens when contracting becomes expensive.

\subsection{The switch from patronage to guarding}

Guarding is another action inside the same extension stage game. A household that hires a thief remains occupied but unguarded, so its own vulnerability is governed by the middle term in \eqref{eq:success-probability}. A household that hires a guard remains occupied and guarded, so the success probability of an incoming elite raid falls to $(1-q_t)\chi_t$. This distinction is what later separates patronage from protection: occupancy by itself no longer makes concentrated wealth safe once predation and retaliation have become selective.

Let elite household $i$ offer a guard job at wage $g$ to poor worker $j$. The worker accepts when the wage weakly exceeds both the freelance option and any noncriminal fallback:
\begin{equation}
g \ge \max\{o_{jt}, \underline{u}_{jt}\},
\label{eq:guard-IR}
\end{equation}
where $\underline{u}_{jt}$ is the best noncriminal outside option. Once accepted, the worker remains at the elite house and does not raid in that period. The employer then receives the expected benefit of avoided elite loss:
\begin{equation}
\pi^{\mathrm{guard}}_{it} = q_t \chi_t \gamma_E w_{it} - g,
\label{eq:guard-payoff}
\end{equation}
where $\gamma_E w_{it}$ is the elite stock that would be exposed in a successful raid. Relative to one more patron-thief contract, guarding is optimal whenever the avoided expected elite loss exceeds the surplus from continued delegated predation:
\begin{equation}
q_t \chi_t \gamma_E w_{it} - g \ge \beta\left(L_t - k - o_{jt}\right).
\label{eq:guard-dominance}
\end{equation}

\begin{theorem}[Endogenous switch to guarding]
\label{thm:guarding}
Suppose guard labor is available at wage $g$ satisfying \eqref{eq:guard-IR} and patronage is initially feasible. Then guarding is optimal for elite agent $i$ whenever
\begin{equation}
w_{it} \ge \widehat{w}^{G}_t \equiv \frac{\beta\left(L_t - k - o_{jt}\right) + g}{q_t \chi_t \gamma_E}.
\label{eq:guard-threshold}
\end{equation}
Moreover, $\widehat{w}^{G}_t$ falls as the prey base among the poor is depleted, because $L_t - o_{jt}$ declines with $\bar{w}^{H}_t$ and $\bar{w}^{P}_t$.
\end{theorem}

\begin{proof}
See Appendix \ref{app:proofs}. Rearranging \eqref{eq:guard-dominance} yields the threshold. Depletion reduces the right-hand side of \eqref{eq:guard-dominance}, lowering the elite wealth level at which guarding becomes privately optimal.
\end{proof}

Guard hiring does not appear immediately because elite wealth is not yet large enough and because patronage already keeps the employer at home while preserving some informal deterrence against attacks on patrons. Guards appear later because wealth concentration raises the loss at stake and because the surplus from one more delegated raid shrinks as poor targets are depleted. The direction of predation remains the same across that transition. During patronage, poor thieves are usually robbing poor households for wealthy principals. Once guarding dominates, part of that same labor pool is reassigned from predation to protection.

\begin{corollary}[Comparative statics of guarding]
\label{cor:guard-statics}
The guarding threshold $\widehat{w}^{G}_t$ in \eqref{eq:guard-threshold} is decreasing in guard effectiveness $q_t$, decreasing in elite exposure $\chi_t$ and elite raid severity $\gamma_E$, increasing in the guard wage $g$, and increasing in the residual surplus from patronage $\beta(L_t-k-o_{jt})$. Guarding therefore becomes more likely when elite wealth is highly exposed and easily defended, and less likely when patronage remains highly profitable or guard labor becomes expensive.
\end{corollary}

\begin{proof}
See Appendix \ref{app:proofs}. The claims follow from direct differentiation of \eqref{eq:guard-threshold} with respect to the relevant parameters.
\end{proof}

Stronger guarding technology or greater elite exposure expands the region in which protection is privately optimal, while stronger patronage surplus or more expensive guard labor contracts that region. Policing emerges from the same strategic environment as patronage.

\subsection{Welfare}

The welfare criterion has to capture inequality and resource dissipation at the same time. An incomplete honest shock can move a society organized around reciprocal corruption toward more unequal and more resource-intensive regimes, while aggregate wealth alone remains uninformative because the benchmark conserves total wealth by construction. The analysis therefore uses discounted utilitarian welfare with explicit resource costs:
\begin{equation}
W_0 = \mathbb{E}_0 \sum_{t=0}^{\infty} \delta^t \mathcal{W}_t,
\qquad
\mathcal{W}_t = \sum_{i=1}^{N} u(w_{it}) - c_T n_t^{S} - k n_t^{H} - g n_t^{G} - \Psi_t,
\label{eq:welfare}
\end{equation}
where $u$ is strictly increasing and strictly concave, $n_t^{S}$ is the number of personal thieves, $n_t^{H}$ is the number of active patron-thief contracts, $n_t^{G}$ is the number of guards, and $\Psi_t$ denotes expected punishment losses.

\begin{proposition}[Redistribution, inequality, and resource cost]
\label{prop:welfare}
Suppose aggregate wealth $\sum_i w_{it}$ is fixed at date $t$ and $u$ is concave. Then any mean preserving spread of the wealth distribution weakly lowers $\sum_i u(w_{it})$. If, in addition, two regimes with the same aggregate wealth differ only in that one has weakly greater inequality and strictly greater values of $n_t^{S}$, $n_t^{H}$, $n_t^{G}$, or $\Psi_t$, then the latter regime yields strictly lower period welfare $\mathcal{W}_t$.
\end{proposition}

\begin{proof}
See Appendix \ref{app:proofs}. The first claim is Jensen's inequality. The second follows because the cost terms enter linearly and negatively in \eqref{eq:welfare}.
\end{proof}

A balanced corruption regime can already be normatively poor if theft effort and punishment risk consume resources, even when the distribution remains equal. Once the honest perturbation generates inequality and induces contracting and guarding costs, welfare can fall even if total wealth is approximately conserved in the short run. The relevant object is therefore distribution plus risk plus deadweight cost, and not redistribution alone.

\begin{corollary}[Guarding may improve welfare relative to patronage]
\label{cor:guard-welfare}
Fix the wealth distribution at date $t$. If adopting guarding lowers expected theft effort, contracting activity, and punishment losses by more than the resulting guard wage bill, then guarding raises $\mathcal{W}_t$ relative to patronage even though it need not restore the welfare of the original balanced benchmark.
\end{corollary}

\begin{proof}
See Appendix \ref{app:proofs}. The difference in period welfare is the reduction in theft-related costs minus the new guarding expenditure.
\end{proof}

Organized enforcement can be privately attractive for elites and even socially preferable to active patronage on a utilitarian measure, while still leaving society worse off than the initial equal benchmark once one accounts for the path by which elite wealth was accumulated. The model therefore leaves room for a limited welfare improvement from policing relative to patronage without mistaking that improvement for a return to egalitarian order.

A balanced theft cycle is a coherent benchmark in the unit transfer stage game. A single permanent honest perturbation breaks that benchmark in the repeated game when wealth-dependent exposure is strong enough. After symmetry breaks, proportional theft together with weak poor outside options makes patronage feasible and profitable. Later, depletion of the prey base and concentration of elite wealth make guarding privately optimal. The analytical results identify those margins, and the simulations trace their timing, prevalence, and welfare consequences.
