\section{Simulation Design}
\label{sec:simulation}

The analytical model identifies the margins that matter, but it does not determine a unique stochastic path once many heterogeneous agents interact. After the first honest night, wealth dispersion changes occupational choices, occupational choices change target vulnerability, patronage and guarding reshape the pool of available workers, and the resulting history feeds back into later wealth. A closed form characterization of all continuation paths would require a large stochastic dynamic game with endogenous role assignment. The simulations translate the analytical objects from Section \ref{sec:model} into an explicit transition process and then measure the timing, prevalence, and comparative statics of the resulting regimes.

\subsection{State variables, timing, and implementation}

The simulation state in period $t$ contains four elements: the wealth vector $w_t$, the role vector $r_t$, the contract network among patrons and thieves, and the guard network among elites and guards. Roles take values in $\{\text{honest}, \text{stay}, \text{freelance thief}, \text{hired thief}, \text{patron}, \text{guard}, \text{elite}\}$, with the understanding that an elite is also a patron or protected household when the relevant threshold is met. The state is updated according to the same period timing used in the theory. Wealth is observed first. Rich and elite sets are then classified by threshold rules that use current mean wealth. Rich agents compare the payoffs from abstention, personal theft, and patronage. Elite agents compare patronage to guarding using the threshold in \eqref{eq:guard-threshold}. Poor agents decide whether to accept theft or guard employment according to their participation constraints. Targets are then selected, raids occur, payments are realized, and the next-period state is recorded.

Deterministic code handles the exact event accounting required by the first honest night, while stochastic code governs the richer transition once matching, targeting, and enforcement interact. The baseline seed is fixed at 613, every sweep uses reproducible seed offsets tied to the parameter grid, and every displayed result is generated from the same underlying pipeline without hand-edited intermediates.

By aligning all paths around the first honest night, event time separates the initial perturbation from later adaptation. Many transition models report only long run outcomes, which makes it difficult to distinguish the direct consequence of an initial perturbation from the later consequences of endogenous adjustment. Here the first honest night is isolated in its own deterministic routine, and the later stochastic dynamics are aligned relative to that event. Parameter changes can therefore be read through their channel: amplification of the first local asymmetry, a change in the profitability of contracting, or a change in the private value of protection.

\subsection{Mapping the theory into simulation rules}

To keep the logic of Theorem \ref{thm:no-stationary-one-honest} fully visible in event time, the benchmark simulation stays very close to the theorem itself. Agents begin on a balanced cycle in the unit transfer environment, one agent is permanently honest, and on the first night the honest agent stays home while the original cycle is otherwise followed exactly. This reproduces the accounting in Lemma \ref{lem:first-night}. In later periods, agents compare their wealth to the threshold in \eqref{eq:threshold}. If wealth lies below the threshold, they continue personal theft along the cycle; if wealth exceeds the threshold, they stay home. No contracting, guarding, or stochastic targeting is yet present, so any transition away from the balanced cycle can be attributed directly to the honest perturbation.

Once proportional transfers are introduced, role thresholds are defined relative to current mean wealth. An agent becomes rich when his wealth exceeds $\theta_R \bar{w}_t$, where $\bar{w}_t$ is the cross-sectional mean, and becomes elite when wealth exceeds the higher threshold $\theta_E \bar{w}_t$. These thresholds are reduced form devices that translate the payoff comparisons of Section \ref{sec:model} into a finite-agent simulation. Rich agents are those for whom the exposure term $\rho w_{it}$ makes personal theft relatively unattractive, while elites are those whose stock of wealth makes guarding salient. In the baseline calibration, the thresholds preserve event time separation between the first honest shock and later institution building while still allowing the patronage regime to emerge within the simulation horizon.

Rich agents are sorted by wealth, and candidate thieves are drawn from the poorest available agents who are neither permanently honest nor already employed. A candidate contract is formed only if the feasibility condition in Lemma \ref{lem:contract-feasibility} holds with current state variables. When a contract is feasible, the simulation applies the bargaining rule from \eqref{eq:bargaining-rule}: the thief receives his outside option plus the share of the residual surplus not captured by the patron, and the patron receives the rest. The thief then targets households in the high value poor set $H_t$, while freelancers target vulnerable households by expected loot. This distinction matters because it is what turns patronage into a strict improvement over freelancing for the patron. Delegated theft is more selective, and its returns are split through a contract that reflects the poor worker outside option.

The targeting rule also resolves a confusion that often arises in informal retellings of the story by Calvino. In the patronage regime, poor thieves typically rob poor households for rich patrons. That is, the physical direction of theft is still largely horizontal among the nonelite, but the financial direction of net transfer is vertical because part of the loot is remitted upward. The simulation keeps this distinction explicit by defining a candidate target set that excludes the patron himself and other currently protected elite households. The house of the patron is occupied and therefore hard to raid, and a hired thief who turns immediately on the employer would be leaving the intended regime. Such betrayals do occur later through the residual elite-risk term $\chi_t$, but they are not the normal action in the patronage phase.

Elites then face a second occupational comparison, namely whether the expected surplus from another patron-thief contract is smaller than the expected value of avoided elite loss from a guard. If inequality has risen and the prey base among the poor has thinned, the guarding condition in \eqref{eq:guard-threshold} will be satisfied for some elite households. Guards are again recruited from the poorest available agents, but only if the guard wage exceeds their best alternative. Once hired, a guard reduces the success probability of raids on the protected house and introduces a punishment probability if an attempted elite raid fails. This punishment channel matters for two reasons. It creates a cost wedge between attacking guarded and unguarded households, and it gives the simulation a concrete mechanism by which formal protection suppresses open theft without eliminating the underlying population of agents willing to steal.

Across the simulation, chronology runs through three stages that match the theory. The first stage is the deterministic first night perturbation in the unit transfer benchmark. The second is the patronage stage in the proportional transfer environment, where class differentiation and delegated predation emerge. The third is the enforcement stage, where some elites replace delegated theft with guards and punishment. Endogenous threshold crossings hard-code the chronology, so calendar dates do not drive the switch and the timing of the transition becomes an outcome determined by wealth dispersion, target depletion, outside options, and enforcement effectiveness.

\subsection{Baseline calibration}

The baseline calibration aims at plausibility within the internal units of the model and is not empirically estimated from a specific dataset. The number of agents is set to $N=60$, which is large enough to produce meaningful cross-sectional dynamics and small enough to make network visualizations interpretable. Initial wealth is normalized to one for every agent. The proportional theft intensity is $\gamma=0.35$, so a successful raid extracts a substantial but not total fraction of the victim wealth. The rich threshold is $\theta_R = 1.15$ times current mean wealth, while the elite threshold is $\theta_E = 2.10$ times current mean wealth. These values imply that early windfalls matter but do not instantly produce an elite stratum.

The remaining parameters pin down the costs and comparative advantages of the occupational choices. The cost of theft effort is $c_T = 0.02$. The contracting cost is $k = 0.03$. The baseline guard wage is $g = 0.08$. The target thief share under a contract is $\alpha = 0.45$, while the fixed wage component is calibrated at $s = 0.03$. The success probability of raiding an occupied but unguarded household is low but positive, and the success probability against a guarded elite household is lower still. Punishment after a failed guarded raid is likely but not certain. In words, the calibration assumes a society in which unprotected wealth is vulnerable, protection is costly, and the poor can be induced to work as thieves or guards because their outside options are weak. Table \ref{tab:parameters} reports the core baseline values used in the main figures.

\begin{table}[tbp]
    \centering
    \caption{Baseline simulation parameters. The values define the benchmark environment used in the main figures, while the parameter sweeps vary the key margins systematically around this baseline.}
    \label{tab:parameters}
    \input{tables/table_parameters.tex}
\end{table}

Parameters that govern the first night logic are chosen to mirror the benchmark theory rather than to generate dramatic inequality mechanically. Parameters that govern patronage and enforcement are chosen so that each regime can emerge for reasons consistent with the propositions in Section \ref{sec:model}. For example, the baseline outside option of poor thieves is set low enough that patronage can be strictly profitable, but not so low that every rich agent hires a thief immediately. Likewise, guard effectiveness is set high enough that protection can dominate when wealth is sufficiently concentrated, but not so high that guards appear before a meaningful patronage phase has existed.

The baseline path is kept moderate: inequality rises without exploding instantly, guards appear only after a visible patronage phase, and collapse remains possible without becoming inevitable. A more aggressive calibration would produce sharper pictures, but it would also obscure whether the mechanisms emerge from the stated thresholds or from extreme parameter values.

\subsection{Parameter sweeps and regime classification}

Systematic parameter sweeps then vary the margins highlighted in the theory. Theft intensity $\gamma$ is varied to measure how sensitive class formation is to the fraction of wealth that can be appropriated in a single raid. Exposure risk is varied through the success probability of attacks on occupied but unguarded households, which determines how costly it is for a rich household to rely on occupancy alone. Contracting cost $k$ and patron bargaining power $\beta$ are varied to map the feasibility and profitability of delegation. Guard effectiveness and punishment severity are varied to locate the transition from patronage to protection. The outside option of poor thieves is varied through the congestion parameter $\lambda_t$ and the availability of guard employment. Finally, the size of the initial honest set is varied so that the simulations can distinguish a single local perturbation from a broader coordinated honesty shock.

Each period in each simulation path is classified into one of six regimes: balanced cycle, transition, patronage, guard regime, coexistence, or collapse. The balanced-cycle label is reserved for periods in which almost all agents continue the reciprocal pattern and inequality remains negligible. Transition denotes periods after the first honest disturbance but before patronage or guarding has become dominant. Patronage denotes periods with active patron-thief contracts and no meaningful guard presence. Guard regime denotes periods in which protected elites and guards are established at scale. Coexistence denotes periods in which patronage and guarding coexist in nontrivial numbers. Collapse denotes periods in which the poor prey base is so depleted that predatory and protective roles no longer reproduce the earlier structure. These labels are not rhetorical. They are deterministic functions of role counts, inequality thresholds, and guard presence stored in the simulation output. The phase diagram in Section \ref{sec:results} is built directly from these classifications.

Fixed coding rules determine regime classification. Regimes are not inferred from unsupervised clustering, and paths are not relabeled ex post to fit the figures. The same classification rule is imposed across the entire parameter grid, which is necessary if a phase diagram is to carry comparative meaning.

\subsection{Recorded outcomes}

For each path the code stores agent level wealth, role histories, employer-worker links, target links, guard assignments, realized and expected welfare components, inequality measures, and first entry times into patronage and guard states. This makes it possible to construct several types of figures that the theory alone cannot provide: event time wealth trajectories by eventual role, distributional heatmaps and quantile ribbons, Lorenz curves at key dates, phase diagrams across parameter pairs, and network snapshots of theft, patronage, and guarding. The same stored objects support the tabulated summaries of first patron entry, first guard entry, final inequality, and dominant long run regime.

Recorded outcomes answer a limited set of analytical questions tied to the theory: how the first honest perturbation propagates through the wealth distribution, when delegated predation becomes organizationally stable, and when organized protection replaces or disciplines predation. Event time plots establish the timeline. Distributional plots establish stratification. Comparative statics show that patronage and guarding do not depend on one baseline path. Welfare decompositions track what is lost and what, at times, is partially recovered as regimes change.
